---
title: "preprocessingDMRs.Rmd"
date: "2025-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# GitHub-ready relative paths. Run this script from the repository root.
project_dir <- getwd()
input_dir <- file.path(project_dir, "input")
output_dir <- file.path(project_dir, "output")
plots_dir <- file.path(output_dir, "plots")
functions_dir <- file.path(project_dir, "functions")

dir.create(input_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(plots_dir, showWarnings = FALSE, recursive = TRUE)

```

## R Markdown

This script is about: 
The DMR data provided has DMR chromosomal location / seqnames, start and end.
This column needs further processing and mapping to ensembl ids of hg19 assembly. 
Region of interest is 2500bp on both endsof the dmr region and extracting genes in the extended area using hg19.gtf. 
Then fetching corresponding gene ids

PART 1: Preprocessing incidentHTN DMRs
PART 2: Preprocessing deltaSBP DMRs

```{r Libraries}

library(tidyverse)
library(GenomicFeatures)
library(AnnotationDbi)
library(org.Hs.eg.db)

#BiocManager::install("TxDb.Hsapiens.UCSC.hg19.knownGene")
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
# setwd() removed for GitHub release; run from repository root.
```


``` {r Reading the DMR data table}

dmr_df <-  read.table(file.path(input_dir, "DMRS_CKD_T2_withT1_matchedcontrols_bumpscutoff_0.02_boots_500.txt"),sep = "\t",  header = TRUE) 

# remove confusing/unwanted columns 
dmr_df <- dmr_df[, !(colnames(dmr_df) %in% c("Gene_Name", "Gene_start", "Gene_end"))]
 
```

Extension to 
2 megabases (Mb)

```{r setup, include=FALSE}

library(GenomicRanges)
library(rtracklayer)

# Load your DMR dataframe
# dmr_df <- read.csv("dmr_file.csv")  # Replace with your actual file path
# Assuming the DMR dataframe has columns: chr, start, end
 
# Extend the DMR regions by 2500 bp on both sides
dmr_df$start_extended <- pmax(0, dmr_df$start - 2000000)
dmr_df$end_extended <- dmr_df$end + 2000000


# Create a GenomicRanges object for DMRs with extended regions
dmr_gr <- GRanges(seqnames = dmr_df$chr,
                  ranges = IRanges(start = dmr_df$start_extended, end = dmr_df$end_extended))

# Load the gene annotation file (GTF format for hg19)
genes_gtf <- import(file.path(input_dir, "gencode.v19.annotation.gtf"))  # Replace with your actual file path

# Filter for gene features (assuming 'gene' is the feature type in the GTF file)
genes_gr <- genes_gtf[genes_gtf$type == "gene"]

# Extract Ensembl Gene IDs (assumed to be in the "gene_id" column of the GTF file)
# Ensembl gene IDs are typically in the "gene_id" attribute in the GTF file
gene_ids <- mcols(genes_gr)$gene_id
gene_names <- mcols(genes_gr)$gene_name

# Find overlaps between the extended DMRs and genes
overlap <- findOverlaps(dmr_gr, genes_gr)

# Get the DMRs and genes that overlap
dmr_overlaps <- data.frame(
  dmr_chr = seqnames(dmr_gr)[queryHits(overlap)],
  dmr_start = start(dmr_gr)[queryHits(overlap)],
  dmr_end = end(dmr_gr)[queryHits(overlap)],
  gene_chr = seqnames(genes_gr)[subjectHits(overlap)],
  gene_start = start(genes_gr)[subjectHits(overlap)],
  gene_end = end(genes_gr)[subjectHits(overlap)],
  gene_name = gene_names[subjectHits(overlap)],
  gene_id = gene_ids[subjectHits(overlap)]  # Extract Ensembl Gene ID
)

# Merge with the other columns from the original dmr_df (use queryHits to link DMR rows)
# Merge with the other columns from the original dmr_df using all three columns: chr, start, and end
dmr_overlaps$dmr_chr <- as.character(dmr_overlaps$dmr_chr)

# Perform a left join using dplyr::left_join to have all other columns 
dmr_overlaps_full <- dmr_overlaps %>%
  left_join(dmr_df, by = c("dmr_chr" = "chr", "dmr_start" = "start_extended", "dmr_end" = "end_extended"))

dmr_overlaps_full$chr <- dmr_overlaps_full$dmr_chr  # keeping a copy for dmr
 
# Rename columns in the dataframe
colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "dmr_start"] <- "dmr_extended_start"
colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "dmr_end"] <- "dmr_extended_end"
colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "start"] <- "dmr_start"
colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "end"] <- "dmr_end"

# Check the updated column names
colnames(dmr_overlaps_full)

                           
# Optional: If you want to output to a CSV file
write.csv(dmr_overlaps_full, file.path(output_dir, "input_2MBpreprocessed_dmr_geneoverlaps.csv"), row.names = FALSE)

# View the result
head(dmr_overlaps_full)

```