---
title: "CorrelationPlot_eQTM.Rmd"
output: html_document
date: "01-08-2025"
Author: Dr. Manasa Kalya
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# GitHub-ready relative paths. Run this script from the repository root.
project_dir <- getwd()
input_dir <- file.path(project_dir, "input")
output_dir <- file.path(project_dir, "output")
plots_dir <- file.path(output_dir, "plots")
functions_dir <- file.path(project_dir, "functions")

dir.create(input_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(plots_dir, showWarnings = FALSE, recursive = TRUE)

```

## R Markdown

This RMD file is to create correlationPlots for eQTM outputs. 
CKD binary Cohort - samples and controls 
1. eqtm is run for whole cohort of 31 controls and 19 samples
2. eQTM correlations within class of controls 
3. eqtm correlations within class of samples 

All these are to be depicted in the final - publication ready plots !

```{r read RDS file}

load(file.path(output_dir, "output_eQTM_CKD_binary.RData"))

#reading meta_info file 
all_meta <- read.csv(file.path(input_dir, "2025.09.29.RNA_CKD_Muhau_Pheno_Timepoint2.csv"))

```

## Including Plots

You can also embed plots, for example:

```{r setting env, echo=FALSE}
# Assuming your function and libraries are loaded

# source(file.path(functions_dir, "eQTM_plot.R"))

library(biomaRt)
library(grid)
library(Gviz)
library(tidyr)
library(ggpubr)
library(gridExtra)
library(SummarizedExperiment)

meth_data <- assays(eqtm.hg19.ckd$dmrs_se)$meth_summarized
expr_data <- assays(eqtm.hg19.ckd$dmrs_se)$expr

# Assuming all_meta_info$R1R2caseControl is a vector where rownames match colnames of expr_data and meth_data
# Find the matching columns between expr_data and R2_RNA_number
matching_columns <- intersect(colnames(expr_data), all_meta$R2_RNA_number)
print(matching_columns)

# Correct way to subset based on matching column names
expr_groups <- all_meta$defCKD_R2[match(colnames(expr_data), all_meta$R2_RNA_number)]
meth_groups <- all_meta$defCKD_R2[match(colnames(meth_data), all_meta$R2_RNA_number)]

# Convert the 0 and 1 values in R1R2caseControl to "control" and "sample"
expr_groups <- ifelse(expr_groups == 0, "control", "sample")
meth_groups <- ifelse(meth_groups == 0, "control", "sample")

# Create united_groups by combining meth_groups and expr_groups
united_groups <- paste(meth_groups, expr_groups, sep = "_")

#dmrs_se_2 <- eqtm.hg38$dmrs_se 
anno_gr <- rowRanges(eqtm.hg19.ckd$dmrs_se)

```

######################################################################################################
####################### FINAL PLOTTING ##############################################


## Lets plot for ZFP5 > ENSG00000204644
index for this gene in eqtm data object will be stored in (index) column 
For ZFP5, index = 1717


sum(expr_values == 0)
[1] 4 
so, there are samples with 0 value in expression before the log-transformation and these appear as dots on the 0 value table 

full : -0.775930742	-0.865611988	-0.680088124	0.001
controls : -0.817873038	-0.892736949	-0.714671319	6	0.001
samples : -0.754858452	-0.882062577	-0.342016078	6	0.001


```{r Final Plot FOR ZFP57 > with manual stats for main and intracalss eqtm, echo=FALSE}

# Load required libraries
library(ggplot2)

# Define the specific index (character format)
#index <- "33"
index = "1717"  #ZFP57

# Extract the corresponding gene name
gene_name <- rowData(eqtm.hg19.ckd$dmrs_se)$geneid[rowData(eqtm.hg19.ckd$dmrs_se)$index == index]


# Find the correct row index based on the extracted index
matched_index <- which(rowData(eqtm.hg19.ckd$dmrs_se)$index == index)

# Extract expression & methylation values
expr_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$expr[matched_index, ])
meth_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$meth_summarized[matched_index, ])

# **Manually define correlation statistics**
# --- Main dataset (overall) ---
cor_coef <- -0.775
ci_lower <- -0.865
ci_upper <- -0.680
pval <- 0.001

# --- Controls only ---
control_cor_coef <- -0.817
control_ci_lower <- -0.892
control_ci_upper <- -0.714
control_pval <- 0.001

# --- Samples only ---
sample_cor_coef <- -0.754
sample_ci_lower <- -0.882
sample_ci_upper <- -0.342
sample_pval <- 0.001

# Create dataframe for plotting
plot_df <- data.frame(
  expr = log2(expr_values + 1.1),  # Log2-transform expression values
  meth = meth_values,
  Sample = colnames(eqtm.hg19.ckd$dmrs_se),
  Group = united_groups  # Use the precomputed sample/control groups
)

# Define correlation annotation text (overall)
cor_text <- paste0( 
                   "\nOverall: r = ", cor_coef, 
                   "\n95% CI: [", ci_lower, ", ", ci_upper, "]",
                   "\nP-value = ", pval)

# Define correlation annotation for control group only
control_cor_text <- paste0("controls: r = ", control_cor_coef, 
                           "\n95% CI: [", control_ci_lower, ", ", control_ci_upper, "]",
                           "\nP-value = ", control_pval)

# Define correlation annotation for sample group only
sample_cor_text <- paste0("case: r = ", sample_cor_coef, 
                          "\n95% CI: [", sample_ci_lower, ", ", sample_ci_upper, "]",
                          "\nP-value = ", sample_pval)

# Rename "control_control" to "controls" and "sample_sample" to "Case"
plot_df$Group <- gsub("control_control", "Control", plot_df$Group)
plot_df$Group <- gsub("sample_sample", "Case", plot_df$Group)

  
ggplot(plot_df, aes(x = expr, y = meth)) +
  
  # Scatter points with respective group colors
  geom_point(aes(fill = Group), shape = 21, size = 4, color = "black", alpha = 0.8) +  

  # Overall regression line (for all data points) - solid black
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "black", linetype = "solid") +  

  # Regression line for Controls - dashed red
  geom_smooth(data = subset(plot_df, Group == "Control"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "red", fill = "#F8766D", linetype = "dashed") +  

  # Regression line for Samples - dashed blue
  geom_smooth(data = subset(plot_df, Group == "Case"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "blue", fill = "#00BFC4", linetype = "dashed") +  
 # Regression line
  theme_bw() +
  xlab(expression(paste("Log2 Gene expression"))) +  
  ylab("median β-value of DMR") +
  ggtitle(paste("Correlation Plot for Gene: ZFP57 (", gene_name, ")")) +  

  # Main correlation annotation (top-right)
  annotate("text", x = min(plot_df$expr) , y = max(plot_df$meth) * 0.6,
           label = cor_text, hjust = 0, size = 3, fontface = "bold", color = "black") +  

  # Control-specific annotation (bottom-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.5, 
           label = control_cor_text, hjust = 0, size = 3, fontface = "bold", color = "red") +  

  # Sample-specific annotation (middle-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.4 ,
           label = sample_cor_text, hjust = 0, size = 3, fontface = "bold", color = "#00BFC4") +  

  scale_fill_manual(values = c("Control" = "#F8766D", "Case" = "#00BFC4",
                               "control_sample" = "#C77CFF", "sample_control" = "#7CAE00")) +  # Custom colors for group combinations
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

# Save the updated plot
ggsave(file.path(plots_dir, "ZFP57(ENSG00000204644)_correlationplot.png"), dpi = 600, width = 10, height = 6)

```

Gene HLA-DRB5 
ENSG00000198502 - 
full_data -0.800061467	-0.865262769	-0.635535749	0.001
controls -0.77577863	-0.890124731	-0.525986199	3	0.001
samples -0.884145108	-0.950836397	-0.742735809	3	0.001


index > 1321


```{r Final Plot FOR HLA-DRB5 gene ENSG00000198502 > with manual stats for main and intracalss eqtm, echo=FALSE}

# Load required libraries
library(ggplot2)

# Define the specific index (character format)
#index <- "33"
index = "1321"  #HLA-DRB5

# Extract the corresponding gene name
gene_name <- rowData(eqtm.hg19.ckd$dmrs_se)$geneid[rowData(eqtm.hg19.ckd$dmrs_se)$index == index]


# Find the correct row index based on the extracted index
matched_index <- which(rowData(eqtm.hg19.ckd$dmrs_se)$index == index)

# Extract expression & methylation values
expr_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$expr[matched_index, ])
meth_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$meth_summarized[matched_index, ])

# **Manually define correlation statistics**
# --- Main dataset (overall) ---
cor_coef <- -0.800
ci_lower <- 	-0.865
ci_upper <- -0.635
pval <- 0.001

# --- Controls only ---
control_cor_coef <- -0.775
control_ci_lower <- 	-0.890
control_ci_upper <- 	-0.525
control_pval <- 0.001

# --- Samples only ---
sample_cor_coef <- -0.884
sample_ci_lower <--0.950
sample_ci_upper <- -0.742
sample_pval <- 0.001

# Create dataframe for plotting
plot_df <- data.frame(
  expr = log2(expr_values + 1),  # Log2-transform expression values
  meth = meth_values,
  Sample = colnames(eqtm.hg19.ckd$dmrs_se),
  Group = united_groups  # Use the precomputed sample/control groups
)

# Define correlation annotation text (overall)
cor_text <- paste0( 
                   "\nOverall: r = ", cor_coef, 
                   "\n95% CI: [", ci_lower, ", ", ci_upper, "]",
                   "\nP-value = ", pval)

# Define correlation annotation for control group only
control_cor_text <- paste0("controls: r = ", control_cor_coef, 
                           "\n95% CI: [", control_ci_lower, ", ", control_ci_upper, "]",
                           "\nP-value = ", control_pval)

# Define correlation annotation for sample group only
sample_cor_text <- paste0("case: r = ", sample_cor_coef, 
                          "\n95% CI: [", sample_ci_lower, ", ", sample_ci_upper, "]",
                          "\nP-value = ", sample_pval)

# Rename "control_control" to "controls" and "sample_sample" to "Case"
plot_df$Group <- gsub("control_control", "Control", plot_df$Group)
plot_df$Group <- gsub("sample_sample", "Case", plot_df$Group)

  
ggplot(plot_df, aes(x = expr, y = meth)) +
  
  # Scatter points with respective group colors
  geom_point(aes(fill = Group), shape = 21, size = 4, color = "black", alpha = 0.8) +  

  # Overall regression line (for all data points) - solid black
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "black", linetype = "solid") +  

  # Regression line for Controls - dashed red
  geom_smooth(data = subset(plot_df, Group == "Control"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "red", fill = "#F8766D", linetype = "dashed") +  

  # Regression line for Samples - dashed blue
  geom_smooth(data = subset(plot_df, Group == "Case"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "blue", fill = "#00BFC4", linetype = "dashed") +  
 # Regression line
  theme_bw() +
  xlab(expression(paste("Log2 Gene expression"))) +  
  ylab("median β-value of DMR") +
  ggtitle(paste("Correlation Plot for Gene: HLA-DRB5 (", gene_name, ")" )) +  

  # Main correlation annotation (top-right)
  annotate("text", x = min(plot_df$expr) , y = max(plot_df$meth) * 0.6,
           label = cor_text, hjust = 0, size = 3, fontface = "bold", color = "black") +  

  # Control-specific annotation (bottom-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.3, 
           label = control_cor_text, hjust = 0, size = 3, fontface = "bold", color = "red") +  

  # Sample-specific annotation (middle-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.1 ,
           label = sample_cor_text, hjust = 0, size = 3, fontface = "bold", color = "#00BFC4") +  

  scale_fill_manual(values = c("Control" = "#F8766D", "Case" = "#00BFC4",
                               "control_sample" = "#C77CFF", "sample_control" = "#7CAE00")) +  # Custom colors for group combinations
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

# Save the updated plot
# Save the updated plot
ggsave(file.path(plots_dir, "HLA-DRB5(ENSG00000198502)_correlationplot.png"), dpi = 600, width = 10, height = 6)

```

################

Gene : SLC39A4(ENSG00000147804), index = 7412
full_data -0.607221989	-0.727821952	-0.417571111  0.001
controls -0.539435241	-0.730900721	-0.253520744	3	0.001
samples -0.704320697	-0.871014967	-0.396835306	3	0.001


```{r Final Plot FOR SLC39A4(ENSG00000258484) > with manual stats for main and intracalss eqtm, echo=FALSE}

# Load required libraries
library(ggplot2)

# Define the specific index (character format)
#index <- "33"
index = "7412"  #SLC39A4

# Extract the corresponding gene name
gene_name <- rowData(eqtm.hg19.ckd$dmrs_se)$geneid[rowData(eqtm.hg19.ckd$dmrs_se)$index == index]


# Find the correct row index based on the extracted index
matched_index <- which(rowData(eqtm.hg19.ckd$dmrs_se)$index == index)

# Extract expression & methylation values
expr_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$expr[matched_index, ])
meth_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$meth_summarized[matched_index, ])

# **Manually define correlation statistics**
# --- Main dataset (overall) ---
cor_coef <- -0.607
ci_lower <- -0.727
ci_upper <- -0.417
pval <- 0.001

# --- Controls only ---
control_cor_coef <- -0.539
control_ci_lower <- -0.730
control_ci_upper <- -0.253
control_pval <- 0.001

# --- Samples only ---
sample_cor_coef <- -0.704
sample_ci_lower <- -0.871
sample_ci_upper <--0.396
sample_pval <- 0.001

# Create dataframe for plotting
plot_df <- data.frame(
  expr = log2(expr_values + 1),  # Log2-transform expression values
  meth = meth_values,
  Sample = colnames(eqtm.hg19.ckd$dmrs_se),
  Group = united_groups  # Use the precomputed sample/control groups
)

# Define correlation annotation text (overall)
cor_text <- paste0( 
                   "\nOverall: r = ", cor_coef, 
                   "\n95% CI: [", ci_lower, ", ", ci_upper, "]",
                   "\nP-value = ", pval)

# Define correlation annotation for control group only
control_cor_text <- paste0("controls: r = ", control_cor_coef, 
                           "\n95% CI: [", control_ci_lower, ", ", control_ci_upper, "]",
                           "\nP-value = ", control_pval)

# Define correlation annotation for sample group only
sample_cor_text <- paste0("case: r = ", sample_cor_coef, 
                          "\n95% CI: [", sample_ci_lower, ", ", sample_ci_upper, "]",
                          "\nP-value = ", sample_pval)

# Rename "control_control" to "controls" and "sample_sample" to "Case"
plot_df$Group <- gsub("control_control", "Control", plot_df$Group)
plot_df$Group <- gsub("sample_sample", "Case", plot_df$Group)

  
ggplot(plot_df, aes(x = expr, y = meth)) +
  
  # Scatter points with respective group colors
  geom_point(aes(fill = Group), shape = 21, size = 4, color = "black", alpha = 0.8) +  

  # Overall regression line (for all data points) - solid black
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "black", linetype = "solid") +  

  # Regression line for Controls - dashed red
  geom_smooth(data = subset(plot_df, Group == "Control"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "red", fill = "#F8766D", linetype = "dashed") +  

  # Regression line for Samples - dashed blue
  geom_smooth(data = subset(plot_df, Group == "Case"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "blue", fill = "#00BFC4", linetype = "dashed") +  
 # Regression line
  theme_bw() +
  xlab(expression(paste("Log2 Gene expression"))) +  
  ylab("median β-value of DMR") +
  ggtitle(paste("Correlation Plot for Gene: SLC39A4(", gene_name, ")")) +  

  # Main correlation annotation (top-right)
  annotate("text", x = min(plot_df$expr) , y = max(plot_df$meth) * 0.6,
           label = cor_text, hjust = 0, size = 3, fontface = "bold", color = "black") +  

  # Control-specific annotation (bottom-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.4, 
           label = control_cor_text, hjust = 0, size = 3, fontface = "bold", color = "red") +  

  # Sample-specific annotation (middle-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.2 ,
           label = sample_cor_text, hjust = 0, size = 3, fontface = "bold", color = "#00BFC4") +  

  scale_fill_manual(values = c("Control" = "#F8766D", "Case" = "#00BFC4",
                               "control_sample" = "#C77CFF", "sample_control" = "#7CAE00")) +  # Custom colors for group combinations
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

# Save the updated plot
# Save the updated plot
ggsave(file.path(plots_dir, "SLC39A4(ENSG00000258484)_correlationplot.png"), dpi = 600, width = 10, height = 6)

```

Gene > PCDHB9 (r = 0.51) (ENSG00000177839), index = 4150

Full data 0.511203328	0.198277745	0.789465836	0.004
controls > 0.545892686	0.016111567	0.896154497	4	0.005
samples 0.622971323	0.162448723	0.899972741	4	0.011



```{r Final Plot FOR PCDHB9(ENSG00000177839) > with manual stats for main and intracalss eqtm, echo=FALSE}

# Load required libraries
library(ggplot2)

# Define the specific index (character format)
#index <- "33"
index = "4150"  #PCDHB9

# Extract the corresponding gene name
gene_name <- rowData(eqtm.hg19.ckd$dmrs_se)$geneid[rowData(eqtm.hg19.ckd$dmrs_se)$index == index]


# Find the correct row index based on the extracted index
matched_index <- which(rowData(eqtm.hg19.ckd$dmrs_se)$index == index)

# Extract expression & methylation values
expr_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$expr[matched_index, ])
meth_values <- as.vector(assays(eqtm.hg19.ckd$dmrs_se)$meth_summarized[matched_index, ])

# **Manually define correlation statistics**
# --- Main dataset (overall) ---
cor_coef <- 0.511
ci_lower <- 0.198
ci_upper <- 0.789
pval <- 0.004

# --- Controls only ---
control_cor_coef <- 0.545
control_ci_lower <- 0.016
control_ci_upper <- 0.896
control_pval <- 0.005

# --- Samples only ---
sample_cor_coef <- 0.622
sample_ci_lower <- 0.162
sample_ci_upper <- 0.899
sample_pval <- 0.011

# Create dataframe for plotting
plot_df <- data.frame(
  expr = log2(expr_values + 1.1),  # Log2-transform expression values
  meth = meth_values,
  Sample = colnames(eqtm.hg19.ckd$dmrs_se),
  Group = united_groups  # Use the precomputed sample/control groups
)

# Define correlation annotation text (overall)
cor_text <- paste0( 
                   "\nOverall: r = ", cor_coef, 
                   "\n95% CI: [", ci_lower, ", ", ci_upper, "]",
                   "\nP-value = ", pval)

# Define correlation annotation for control group only
control_cor_text <- paste0("controls: r = ", control_cor_coef, 
                           "\n95% CI: [", control_ci_lower, ", ", control_ci_upper, "]",
                           "\nP-value = ", control_pval)

# Define correlation annotation for sample group only
sample_cor_text <- paste0("case: r = ", sample_cor_coef, 
                          "\n95% CI: [", sample_ci_lower, ", ", sample_ci_upper, "]",
                          "\nP-value = ", sample_pval)

# Rename "control_control" to "controls" and "sample_sample" to "Case"
plot_df$Group <- gsub("control_control", "Control", plot_df$Group)
plot_df$Group <- gsub("sample_sample", "Case", plot_df$Group)

  
ggplot(plot_df, aes(x = expr, y = meth)) +
  
  # Scatter points with respective group colors
  geom_point(aes(fill = Group), shape = 21, size = 4, color = "black", alpha = 0.8) +  

  # Overall regression line (for all data points) - solid black
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "black", linetype = "solid") +  

  # Regression line for Controls - dashed red
  geom_smooth(data = subset(plot_df, Group == "Control"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "red", fill = "#F8766D", linetype = "dashed") +  

  # Regression line for Samples - dashed blue
  geom_smooth(data = subset(plot_df, Group == "Case"), method = "lm", formula = y ~ x, 
              aes(fill = Group), se = TRUE, color = "blue", fill = "#00BFC4", linetype = "dashed") +  
 # Regression line
  theme_bw() +
  xlab(expression(paste("Log2 Gene expression"))) +  
  ylab("median β-value of DMR") +
  ggtitle(paste("Correlation Plot for Gene: PCDHB9", gene_name, ")")) +  

  # Main correlation annotation (top-right)
  annotate("text", x = min(plot_df$expr) , y = max(plot_df$meth) * 0.45,
           label = cor_text, hjust = 0, size = 3, fontface = "bold", color = "black") +  

  # Control-specific annotation (bottom-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.35, 
           label = control_cor_text, hjust = 0, size = 3, fontface = "bold", color = "red") +  

  # Sample-specific annotation (middle-left)
  annotate("text", x = min(plot_df$expr), y = max(plot_df$meth) * 0.25 ,
           label = sample_cor_text, hjust = 0, size = 3, fontface = "bold", color = "#00BFC4") +  

  scale_fill_manual(values = c("Control" = "#F8766D", "Case" = "#00BFC4",
                               "control_sample" = "#C77CFF", "sample_control" = "#7CAE00")) +  # Custom colors for group combinations
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

# Save the updated plot
ggsave(file.path(plots_dir, "PCDHB9(ENSG00000177839)_correlationplot.png"), dpi = 600, width = 10, height = 6)

```

