---
title: "RODAM_CKD_eqtm"
output: html_document
date: "01-08-25"
Author : Dr.Manasa Kalya 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# GitHub-ready relative paths. Run this script from the repository root.
project_dir <- getwd()
input_dir <- file.path(project_dir, "input")
output_dir <- file.path(project_dir, "output")
plots_dir <- file.path(output_dir, "plots")
functions_dir <- file.path(project_dir, "functions")

dir.create(input_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(plots_dir, showWarnings = FALSE, recursive = TRUE)

```

## R Markdown

BATCH 1 > Intraclass_Controls 
BATCH 2 > Intraclass_Samples 


```{r Libraries, include=FALSE}
library(dplyr)
library(GenomicRanges)
library(data.table)
library(tidyverse)
library(readr)
# setwd() removed for GitHub release; run from repository root.
```




```{r load sample_info file, include=FALSE}

all_meta <- read.csv(file.path(input_dir, "2025.07.10.RNA_CKD_Muhau_Pheno_Timepoint2.csv"))

meta_no_duplicates <- all_meta[!duplicated(all_meta$RodamID), ] # Check/remove duplicates if any
meta_no_duplicates <- meta_no_duplicates %>% remove_rownames %>% column_to_rownames(var="R2_RNA_number")
meta_no_duplicates$R2_RNA_number <- rownames(meta_no_duplicates)

# Create a named vector mapping 'Basename' to 'R2_RNA_number'
basename_to_rna <- setNames(meta_no_duplicates$R2_RNA_number, meta_no_duplicates$Basename.x)

# Check the mapping
head(basename_to_rna)

```


```{r Samples to subset for intraclass correlation, include=FALSE}

# Extract samples where R1R2SBPControl == 0
intraclass_controls <- rownames(meta_no_duplicates[meta_no_duplicates$defCKD_R2 == 0, ])
intraclass_samples <- rownames(meta_no_duplicates[meta_no_duplicates$defCKD_R2 == 1, ])

```

## eQTM Analysis

This function was setup to perform an expression quantitative trait methylation analysis for methylation and gene expression data obtained from the same samples.
# 1.' @param dmrs_gr A GenomicRanges-derived object where the ranges represent the coordinates of the DMR and whose metadata contains an associated gene identifier.
# 2. ' @param gene_col A column index containing of the dmrs_gr object containing the gene identifiers.
# 3. ' @param meth_data A matrix whose row names contain CpG identifiers and whose values represent the methylation signal. The column names must contain the columns names of the gene expression data.
# 4. ' @param expr_data A matrix whose row names contain gene identifiers and whose values represent the expression signal. The column names must contain the columns names of the methylation data.
# 5. ' @param meth_anno_gr A GenomicRanges-derived object where the ranges represent the coordinates of the assayed CpGs and whose rownames contain an identifier.

This is to process and align epr_data for further analysis: 
There are 2 low-quality RNA-seq samples, which I will be excluding additionally., 
According to RNA-seq FastQ QC, there are 2 samples that are poor and we have decided to remove them from the analysis. 
G2155 & P0874.

```{r liftedover_GEX_data , include=FALSE}

expr_matrix <- fread(file.path(input_dir, "input_hg19_filteredNormalized_GEX_touse.csv"))
#dim(expr_matrix) 
expr_matrix$V1 <- NULL

expr_matrix <- as.data.frame(expr_matrix)

rownames(expr_matrix) <- expr_matrix$hg19_ensembl_id

expr_matrix$hg19_ensembl_id <- NULL

# setwd("<project-relative-path>")


# check for duplicated gene IDs in your expression matrix
# Check how many duplicated gene IDs (row names)
# Remove digits and the period after it from rownames of expr_data
rownames(expr_matrix) <- sub("\\..*", "", rownames(expr_matrix))


#sanity
sum(duplicated(rownames(expr_matrix)))
duplicated_genes <- rownames(expr_matrix)[duplicated(rownames(expr_matrix))]


# additional column for eqtm 
# expr_matrix$geneid <- rownames(expr_matrix)  # Add geneid column from rownames

# Check if expr_data is numeric 
```

# There are no duplicates in rownames(expr_matrix) 
length(rownames(expr_matrix))
[1] 15076
 length(unique(rownames(expr_matrix)))
[1] 15076


```{r @param : junk_1 cleaning expr_data matrix, echo=FALSE}
# 
# expr_matrix <- fread("<project-relative-path>")
# 
# 
# #dim(expr_matrix) 
# expr_matrix$V1 <- NULL
# 
# # Aggregate the data by the Ensembl ID and take the mean across the duplicates
# # Convert to data.table if it's not already
# setDT(expr_matrix)
# 
# # Remove the 'hg38_ensembl_id' column and aggregate by 'hg19_ensembl_id'
# expr_matrix_aggregated <- expr_matrix[, !'hg38_ensembl_id', with = FALSE] %>%
#   .[, lapply(.SD, mean), by = hg19_ensembl_id]
# 
# ##################################################################################
# # Validation; 20,328 rows, unique(hg19_ensembl_id) --> 7318 only. 
# # 1. length(unique(expr_matrix$hg19_ensembl_id))  ... #7318
# # 2. Crosscheck if leftover is fine: # Check for exact matches between hg19_ensembl_id and hg38_ensembl_id
# matches <- expr_matrix$hg19_ensembl_id == expr_matrix$hg38_ensembl_id
# 
# # Count the number of matching IDs
# num_matches <- sum(matches, na.rm = TRUE)
# 
# # Count the number of discrepancies (non-matches)
# num_discrepancies <- sum(!matches, na.rm = TRUE)
# 
# # Display the results
# cat("Number of matches: ", num_matches, "\n")
# cat("Number of discrepancies: ", num_discrepancies, "\n")   #prints NAN, no miss_matches
# 
# # Optionally, show the mismatched rows (for further inspection)
# discrepant_rows <- expr_matrix[!matches, ]  # ZERO
# ############################################################################################
# 
# # Remove rows with missing Ensembl IDs (if any)
# expr_matrix_aggregated_clean <- expr_matrix_aggregated %>%
#   filter(!is.na(hg19_ensembl_id))
# 
# # Set the Ensembl IDs as row names
# expr_matrix_aggregated_df <- as.data.frame(expr_matrix_aggregated_clean)
# 
# rownames(expr_matrix_aggregated_df) <- expr_matrix_aggregated_df$hg19_ensembl_id
# expr_matrix_aggregated_df$hg19_ensembl_id <- NULL
# 
# # View the updated matrix
# #head(expr_matrix_aggregated_df)
# 
# #################################################################################################
# # remove lowQC Samples
# 
# # Remove columns "G2155" and "P0874" using select() from dplyr
# expr_matrix <- expr_matrix_aggregated_df[, !(colnames(expr_matrix_aggregated_df) %in% c("G2155", "P0874"))]
# 
# # View the updated dataframe
# head(expr_matrix)  # 180 samples 

```



```{r @param 1: load meth_data matrix, echo=FALSE}

#beta_matrix <- fread("<project-relative-path>")

beta_matrix <- fread(file.path(input_dir, "betaFunnurm.txt"))
beta_matrix <- as.data.frame(beta_matrix)
rownames(beta_matrix) <-beta_matrix$V1
beta_matrix$V1 <- NULL
  
```


```{r @param 1: processing beta_matrix further to match expr_data, echo=FALSE}

# Replace the column names of beta_matrix with the corresponding R2_RNA_number
colnames(beta_matrix) <- basename_to_rna[colnames(beta_matrix)]

# subset beta_matrix which have R2_RNA_Number in all_meta
beta_matrix_2 <- beta_matrix[, !is.na(colnames(beta_matrix))]

# Check the new column names in beta_matrix_2
head(colnames(beta_matrix_2))
  

############################################################################
## Subset beta_matrix 

beta_matrix_2 <- beta_matrix_2[, !is.na(colnames(beta_matrix_2))] # cpg Identifier column

# Find the common columns between beta_matrix_2 and expr_data
common_columns <- intersect(colnames(beta_matrix_2), colnames(expr_matrix))
# Check the common columns
common_columns

# Subset beta_matrix_2 to only include the common columns
beta_matrix_2_subset <- beta_matrix_2[, common_columns]
expr_matrix <- expr_matrix[,common_columns]

# Check if ProbeID Column exists, if so - remove it
# beta_matrix_2_subset$probeID <- NULL

# Check the result
head(beta_matrix_2_subset)

# Reorder columns in expr_data and beta_matrix_2 to match each other

# Check if both dataframes have the same column names
all(colnames(expr_matrix) %in% colnames(beta_matrix_2_subset))  # Should return TRUE
all(colnames(beta_matrix_2_subset) %in% colnames(expr_matrix))  # Should return TRUE

# Reorder columns of expr_matrix_aggregated_df to match the order of columns in beta_matrix_2
expr_matrix <- expr_matrix[, colnames(beta_matrix_2_subset)]

# Check if the column names of both dataframes are now the same
all(colnames(expr_matrix) == colnames(beta_matrix_2_subset))  # Should return TRUE

expr_matrix <- as.data.frame(expr_matrix)

# Remove digits and the period after it from rownames of expr_data
# rownames(expr_matrix) <- sub("\\..*", "", rownames(expr_matrix))

```

Annotation file should be placed in the `input/` directory.


```{r @param 3: @param meth_anno_gr, Genomic ranges, echo=FALSE}


annotation_data <- fread(file.path(input_dir, "filteredAnnotation.txt"))

annotation_data = as.data.frame(annotation_data)
colnames(annotation_data)
rownames(annotation_data) <- annotation_data$V1

# Extract relevant columns from annotation_data
cpg_annotation <- annotation_data[, c("chr", "pos", "Name")]
head(cpg_annotation)
# Extract relevant columns from annotation_data
cpg_annotation <- annotation_data[, c("chr", "pos", "Name")]
head(cpg_annotation)

# Make sure the row names of beta_matrix match the 'Name' column in annotation_data
# The row names of your beta matrix represent the CpG probe IDs
# Merge the annotation with the beta matrix by matching probe IDs

# Add rownames as a column for both data frames
cpg_annotation$probeID <- rownames(cpg_annotation)

#beta_matrix_2_subset_df <- data.frame(probeID = rownames(beta_matrix_2_subset), 
                                     #beta_values = as.vector(as.numericbeta_matrix_2_subset))

beta_matrix_2_subset$probeID <- rownames(beta_matrix_2_subset) # temporary column

# Merge on 'probeID' after adding rownames as a column
cpg_data <- merge(cpg_annotation, 
                  as.data.frame(beta_matrix_2_subset), 
                  by.x = "probeID", 
                  by.y = "probeID", 
                  all.x = FALSE)

cpg_data$probeID <- NULL # Optionally, remove the 'probeID' column if no longer needed after merge

dim(cpg_data) # Check the dimensions of the merged data


# Check the merged data
head(cpg_data)
# colnames(cpg_data) <- gsub("beta_values\\.", "", colnames(cpg_data))

rownames(cpg_data) <- cpg_data$Name

## STEP 2: Converting to Genomic Ranges 

meth_anno_gr <- GRanges(
  seqnames = cpg_data$chr,  # Chromosome
  ranges = IRanges(start = cpg_data$pos, end = cpg_data$pos),  # Position (start == end for each CpG site)
  probeID = cpg_data$Name,  # Probe ID
  beta_values = cpg_data[, grep("beta", colnames(cpg_data))]  # Beta values (assuming columns contain beta values)
)

# Set the 'names' of the GRanges object to match the rownames of cpg_data
names(meth_anno_gr) <- rownames(cpg_data)

# Check the names of meth_anno_gr
names(meth_anno_gr)

```




```{r @param 1: load DMR data , echo=FALSE}

df = read.csv(file.path(output_dir, "input_2MBpreprocessed_dmr_geneoverlaps.csv"), header = TRUE) 


# Remove digits and the period after it from rownames of expr_data
df$gene_id <- sub("\\..*", "", df$gene_id)

# # Rename specific columns in dmr_df
# colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "dmr_start"] <- "dmr_extended_start"
# colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "dmr_end"] <- "dmr_extended_end"
# colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "start"] <- "dmr_start"
# colnames(dmr_overlaps_full)[colnames(dmr_overlaps_full) == "end"] <- "dmr_end"

# Verify the changes
print(colnames(dmr_overlaps_full))

# Count number of NA values in gene_id column
#sum(is.na(df$gene_id))    #must be 0, otherwise, remove them
# Assuming your data frame is named df
# Remove version number from Ensembl IDs 
#df$gene_id <- sub("\\.\\d+$", "", df$gene_id)

# Assuming your data frame is named df
# Convert the dataframe to a GRanges object
ckd_dmrs <- GRanges(
  seqnames = df$dmr_chr,  # Chromosome
  ranges = IRanges(start = df$dmr_start, end = df$dmr_end),  # Genomic ranges (start and end positions)
  value = df$value,  # Example of additional data (value)
  area = df$area,  # Additional column (area)
  cluster = df$cluster,  # Additional column (cluster)
  indexStart = df$indexStart,  # Example metadata
  indexEnd = df$indexEnd,
  L = df$L,
  clusterL = df$clusterL,
  pvalue = df$p.value,  # P-values as metadata
  fwer = df$fwer,
  pvalueArea = df$p.valueArea,
  fwerArea = df$fwerArea,
  Gene_start = df$gene_start,  # Gene start position
  Gene_end = df$gene_end, 
  Gene_name = df$gene_name, # Gene end position
  geneid =  df$gene_id     # Gene name
)

# Check the resulting GenomicRanges object
head(ckd_dmrs)

```

1. Error in if (const(t, min(1e-08, mean(t, na.rm = TRUE)/1e+06))) { : 
  missing value where TRUE/FALSE needed - 


```{r @param filtering GEX for low SD  , echo=FALSE}

# # Assuming expr_matrix is your original expression matrix
# # Number of samples (columns)
# num_samples <- ncol(expr_matrix)
# 
# # Define the threshold for zero values and low standard deviation
# zero_threshold <- 0.8 * num_samples  # 80% of samples
# sd_threshold <- 0.1  # Standard deviation threshold, adjust as needed
# 
# # Step 1: Identify rows with 80% or more zero values
# rows_with_zeros <- apply(expr_matrix, 1, function(x) sum(x == 0)) >= zero_threshold
# 
# # Step 2: Identify rows with low standard deviation
# rows_with_low_sd <- apply(expr_matrix, 1, sd) <= sd_threshold
# 
# # Step 3: Combine the conditions to filter rows
# rows_to_remove <- rows_with_zeros | rows_with_low_sd
# 
# # Step 4: Subset the expression matrix
# filtered_expr_matrix <- expr_matrix[!rows_to_remove, ]
# 
# # Remove duplicate row names, keeping only the first occurrence
# filtered_expr_matrix <- filtered_expr_matrix[!duplicated(rownames(filtered_expr_matrix)), ]
# 
# # View the filtered expression matrix
# dim(filtered_expr_matrix)
# 
# filtered_expr_matrix_2 <- filtered_expr_matrix[, apply(filtered_expr_matrix, 2, var) > 1]




# Load the matrices
#exp_matrix <- read.csv("exp_matrix.csv", row.names = 1)
#beta_matrix <- read.csv("beta_matrix.csv", row.names = 1)

#check this always
beta_matrix_2_subset$probeID <- NULL

# # Check for constant rows (zero standard deviation)
# constant_rows_expr <- apply(expr_matrix, 1, function(x) sd(x, na.rm = TRUE) == 0)
# constant_rows_meth <- apply(beta_matrix_2_subset, 1, function(x) sd(x, na.rm = TRUE) == 0)
# 
# # Report how many rows have constant values
# sum(constant_rows_expr)  # Genes with no variation in expression
# sum(constant_rows_meth)  # CpGs with no variation in methylation
# 
# expr_matrix <- expr_matrix[!constant_rows_expr, ]
# beta_matrix <- beta_matrix_2_subset[!constant_rows_meth, ]

sum(is.na(expr_matrix))
sum(is.na(beta_matrix_2_subset))

```



```{r @param : junk_2 : flow variance rows  , echo=FALSE}

# # Check for constant or zero variance rows in methylation data
# row_variances_meth <- apply(beta_matrix_2_subset, 1, var)
# zero_variance_rows_meth <- which(row_variances_meth == 0)
# 
# # Check for constant or zero variance rows in expression data
# row_variances_expr <- apply(filtered_expr_matrix, 1, var)
# zero_variance_rows_expr <- which(row_variances_expr == 0)
# 
# # Print out any problematic rows
# print(zero_variance_rows_meth)
# print(zero_variance_rows_expr)

```

For intraclass eQTM Analysis - 
We subset expr_matrix and beta_matrix according to controls and cases using pheno_file 
Batch1 > Controls 
Batch 2 > Cases 


Batch 1 > Controls 


```{r eqtm function, echo=FALSE}

#require("NDlib")
#require(eqtm)

#library(NDlib)

source(file.path(functions_dir, "PHSugg_eQTM_functions_rodam.R"))


library(SummarizedExperiment)
library(foreach)
library(parallel)


# Step 1 : Subset Expr_matrix

expr_matrix_controls <- expr_matrix[, colnames(expr_matrix) %in% intraclass_controls]

# Step 2 > Subset Beta_matrix

# Find the common columns between beta_matrix_2 and expr_data
common_columns <- intersect(colnames(beta_matrix_2_subset), colnames(expr_matrix_controls))
# Check the common columns
common_columns

# Subset beta_matrix_2 to only include the common columns
beta_matrix_2_controls <- beta_matrix_2_subset[, common_columns]
expr_matrix_controls <- expr_matrix[,common_columns]

# Check if ProbeID Column exists, if so - remove it
# beta_matrix_2_subset$probeID <- NULL

# Check the result
head(beta_matrix_2_controls)

# Reorder columns in expr_data and beta_matrix_2 to match each other

# Check if both dataframes have the same column names
all(colnames(expr_matrix_controls) %in% colnames(beta_matrix_2_controls))  # Should return TRUE
all(colnames(beta_matrix_2_controls) %in% colnames(expr_matrix_controls))  # Should return TRUE

# Reorder columns of expr_matrix_aggregated_df to match the order of columns in beta_matrix_2
expr_matrix_controls <- expr_matrix_controls[, colnames(beta_matrix_2_controls)]

# Check if the column names of both dataframes are now the same
all(colnames(expr_matrix_controls) == colnames(beta_matrix_2_controls))  # Should return TRUE

expr_matrix_controls <- as.data.frame(expr_matrix_controls)

##################################################################

# sanity checks
# beta_matrix$probeID <- NULL # ensure this column is not there in final matrix

## 1. Check typeof beta, expr-data - must be dataframes
typeof(expr_matrix_controls)  #list   needs conversion
typeof(beta_matrix_2_controls)   #list 

## 2. Check dimensions, they must be same 
dim(expr_matrix_controls)
dim(beta_matrix_2_controls)

# 3. colnames must match 
all(colnames(expr_matrix_controls) == colnames(beta_matrix_2_controls)) # must be TRUE 

# 4. Check overlapping geneids - dmr$geneids and rownames(expr_matrix)
d <- expr_matrix_controls[(rownames(expr_matrix_controls) %in% df$gene_id),]
print(dim(d))            #46 overlapping

# 5. Ensure "geneid" is the column_name in rodam_dmrs object
ckd_dmrs$geneid


expr_matrix_controls <- as.matrix(expr_matrix_controls)
beta_matrix_2_controls <- as.matrix(beta_matrix_2_controls)


################################################################################
###### RUN eQTM 

# Define the output directory
output_dir <- "<project-relative-path>"

# Ensure the "output" directory exists
if (!dir.exists(output_dir)) {
  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
}

# Define file paths inside "output" directory
output_rdata_file <- file.path(output_dir, "output_eQTM_CKD_binary_intraclass_controls.RData")
output_csv_file <- file.path(output_dir, "output_eQTM_CKD_binary_intraclass_controls.csv")


# Run the eqtm function
eqtm.hg19.ckd <- eqtm(
  dmrs_gr = ckd_dmrs, 
  gene_col = "geneid", 
  meth_data = beta_matrix_2_controls, 
  expr_data = expr_matrix_controls, 
  meth_anno_gr = meth_anno_gr, 
  cor_method = "pearson", 
  alternative = "two.sided", 
  N = 1000, 
  iseed = 531235, 
  ncores = 4,
  verbose = TRUE
)

save(eqtm.hg19.ckd, file = output_rdata_file)

g <- as.data.frame(eqtm.hg19.ckd$dmrs_chrom_df)
new_colnames <- as.character(g["geneid", ])  # Extract the geneid row as a vector
colnames(g) <- new_colnames

# 4. Transpose the dataframe
g_transposed <- t(g)
g_transposed <- as.data.frame(g_transposed)

# Print the updated g_transposed dataframe with overlapping DMRs
head(g_transposed)

write.csv(g_transposed,output_csv_file,row.names = FALSE)

#save(eqtm.hg19.deltaSBP, file = "eqtm.hg19.deltaSBP.RData")

```


# BATCH 2 > eQTM for intraclass_samples 


```{r Batch 2 > intraclass eQTM for cases , include=FALSE}


source(file.path(functions_dir, "PHSugg_eQTM_functions_rodam.R"))


library(SummarizedExperiment)
library(foreach)
library(parallel)


# Step 1 : Subset Expr_matrix

expr_matrix_samples <- expr_matrix[, colnames(expr_matrix) %in% intraclass_samples]

# Step 2 > Subset Beta_matrix

# Find the common columns between beta_matrix_2 and expr_data
common_columns <- intersect(colnames(beta_matrix_2_subset), colnames(expr_matrix_samples))
# Check the common columns
common_columns

# Subset beta_matrix_2 to only include the common columns
beta_matrix_2_samples <- beta_matrix_2_subset[, common_columns]
expr_matrix_samples <- expr_matrix[,common_columns]

# Check if ProbeID Column exists, if so - remove it
# beta_matrix_2_subset$probeID <- NULL

# Check the result
head(beta_matrix_2_samples)

# Reorder columns in expr_data and beta_matrix_2 to match each other

# Check if both dataframes have the same column names
all(colnames(expr_matrix_samples) %in% colnames(beta_matrix_2_samples))  # Should return TRUE
all(colnames(beta_matrix_2_samples) %in% colnames(expr_matrix_samples))  # Should return TRUE

# Reorder columns of expr_matrix_aggregated_df to match the order of columns in beta_matrix_2
expr_matrix_samples <- expr_matrix_samples[, colnames(beta_matrix_2_samples)]

# Check if the column names of both dataframes are now the same
all(colnames(expr_matrix_samples) == colnames(beta_matrix_2_samples))  # Should return TRUE

expr_matrix_samples <- as.data.frame(expr_matrix_samples)

##################################################################

# sanity checks
# beta_matrix$probeID <- NULL # ensure this column is not there in final matrix

## 1. Check typeof beta, expr-data - must be dataframes
typeof(expr_matrix_samples)  #list   needs conversion
typeof(beta_matrix_2_samples)   #list 

## 2. Check dimensions, they must be same 
dim(expr_matrix_samples)
dim(beta_matrix_2_samples)

# 3. colnames must match 
all(colnames(expr_matrix_samples) == colnames(beta_matrix_2_samples)) # must be TRUE 

# 4. Check overlapping geneids - dmr$geneids and rownames(expr_matrix)
d <- expr_matrix_samples[(rownames(expr_matrix_samples) %in% df$gene_id),]
print(dim(d))            #46 overlapping

# 5. Ensure "geneid" is the column_name in rodam_dmrs object
ckd_dmrs$geneid


expr_matrix_samples <- as.matrix(expr_matrix_samples)
beta_matrix_2_samples <- as.matrix(beta_matrix_2_samples)


################################################################################
###### RUN eQTM 

# Define the output directory
output_dir <- "<project-relative-path>"

# Ensure the "output" directory exists
if (!dir.exists(output_dir)) {
  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
}

# Define file paths inside "output" directory
output_rdata_file <- file.path(output_dir, "output_eQTM_CKD_binary_intraclass_samples.RData")
output_csv_file <- file.path(output_dir, "output_eQTM_CKD_binary_intraclass_samples.csv")


# Run the eqtm function
eqtm.hg19.ckd <- eqtm(
  dmrs_gr = ckd_dmrs, 
  gene_col = "geneid", 
  meth_data = beta_matrix_2_samples, 
  expr_data = expr_matrix_samples, 
  meth_anno_gr = meth_anno_gr, 
  cor_method = "pearson", 
  alternative = "two.sided", 
  N = 1000, 
  iseed = 531235, 
  ncores = 4,
  verbose = TRUE
)

save(eqtm.hg19.ckd, file = output_rdata_file)

g <- as.data.frame(eqtm.hg19.ckd$dmrs_chrom_df)
new_colnames <- as.character(g["geneid", ])  # Extract the geneid row as a vector
colnames(g) <- new_colnames

# 4. Transpose the dataframe
g_transposed <- t(g)
g_transposed <- as.data.frame(g_transposed)

# Print the updated g_transposed dataframe with overlapping DMRs
head(g_transposed)   


write.csv(g_transposed,output_csv_file,row.names = FALSE)

#save(eqtm.hg19.deltaSBP, file = "eqtm.hg19.deltaSBP.RData")

```


