---
title: "Output_eQTM_annotation.Rmd"
output: html_document
date: "2025-02-18"
Author > Manasa Kalya 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# GitHub-ready relative paths. Run this script from the repository root.
project_dir <- getwd()
input_dir <- file.path(project_dir, "input")
output_dir <- file.path(project_dir, "output")
plots_dir <- file.path(output_dir, "plots")
functions_dir <- file.path(project_dir, "functions")

dir.create(input_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(plots_dir, showWarnings = FALSE, recursive = TRUE)

# All packages and libraries sit here: 
# Optional: set a custom library path here if needed.


```

## 
From the eQTM pipeline, we have Ensembl geneids correlated to specific DMRS *based on DMR chromosomal location* , corresponding Correlation coefficient, Confidence interval and p_values. 

Now, this snippet works on adding annotation to these genes and DMRS as > 
a. Whether DMR sits in Promoter of the gene 
b. In the genebody 
c. In the exon 
d. Far away enhancer / based on ENCODE project

```{r load necessary libraries, include=FALSE}

library(GenomicRanges)
library(rtracklayer)
#library(ChIPseeker)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)  # Change to hg38 if needed
#BiocManager::install("txdbmaker")
library(txdbmaker)
# BiocManager::install("ChIPseeker")
library(ChIPseeker)

library(ggtree) # needed for ChIPseeker
# setwd() removed for GitHub release; run from repository root.
```



```{r read eQTM output}

dmr_df <- read.csv(file.path(output_dir, "output_eqtm_CKD_binary.csv"))

head(dmr_df)
```

## Including Chromosomal location for genes *hg19

For Differentially Methylated Region (DMR) annotation, the most biologically relevant features are:

Feature	Why Included?
1. Gene	Defines the entire transcriptional unit (useful for gene-body methylation).
2. Exon	Identifies if a DMR is within coding/non-coding sequences.
3. Promoter	Defined as 2kb upstream of the Transcription Start Site (TSS), crucial for gene regulation.

# Why Were Other GTF Features Excluded?

1. Transcript	Too redundant. A single gene can have multiple transcripts, making annotation noisy.
2. CDS (Coding Sequence)	Less relevant because methylation primarily affects transcription, not translation.
3. UTR (Untranslated Regions)	Optional. UTRs regulate translation but are not primary methylation targets.
4. Start_Codon / Stop_Codon	Too localized. Methylation changes are usually broad rather than codon-specific.

Selenocysteine	Extremely rare and specific to certain tRNA pathways, not relevant for general DMR annotation.

###########################

Let us subset / Gene, Exon, promoter, CDS , S

```{r Annotation, echo=FALSE}

# Convert DMR data to GRanges object
dmr_gr <- GRanges(seqnames = dmr_df$seqnames,
                  ranges = IRanges(start = dmr_df$start, end = dmr_df$end), 
                  geneid = dmr_df$geneid, ncps = dmr_df$nCpGs, cor_coef = dmr_df$cor_coef, 
                  CI95_lower = dmr_df$CI95_lower, CI95_upper = dmr_df$CI95_upper, cor_pvalue = dmr_df$pvals)

# Load Gene Annotation (GTF file)
gtf_file <- file.path(input_dir, "gencode.v19.annotation.gtf")  # Change to your GTF file
gtf_data <- import(gtf_file)

# Extract relevant gene features from GTF
gene_regions <- subset(gtf_data, type == "gene")
transcript_regions <- subset(gtf_data, type == "transcript")
exon_regions <- subset(gtf_data, type == "exon")
cds_regions <- subset(gtf_data, type == "CDS")
utr_regions <- subset(gtf_data, type == "UTR")
start_codon_regions <- subset(gtf_data, type == "start_codon")
stop_codon_regions <- subset(gtf_data, type == "stop_codon")
selenocysteine_regions <- subset(gtf_data, type == "Selenocysteine")  # Rare but included

#  Define introns properly using GenomicFeatures**
txdb <- makeTxDbFromGFF(gtf_file)  # Convert GTF to TxDb format
introns_by_tx <- intronsByTranscript(txdb, use.names = TRUE)  # Extract introns
intron_regions <- unlist(introns_by_tx)  # Convert list to GRanges

# Define promoter regions (2kb upstream of TSS)
promoter_regions <- promoters(gene_regions, upstream = 2000, downstream = 500)

# Combine all feature annotations
all_features <- c(gene_regions, transcript_regions, exon_regions, intron_regions, 
                  cds_regions, utr_regions, start_codon_regions, stop_codon_regions, 
                  selenocysteine_regions, promoter_regions)

seqlevelsStyle(dmr_gr) <- "UCSC"  # Converts to "chr1" format
seqlevelsStyle(txdb) <- "UCSC"  # Ensure TxDb matches "chr1" format

# Extract gene names from the GTF data
gene_info <- data.frame(
  geneid = gene_regions$gene_id,  # Extract gene ID
  gene_name = gene_regions$gene_name  # Extract gene name
)

gene_info$geneid <- sub("\\..*", "", gene_info$geneid)   #remove verson numbers


# Annotate DMRs using ChIPseeker
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene  # Adjust to hg38 if needed
dmr_annotation <- annotatePeak(dmr_gr, TxDb = txdb, tssRegion = c(-2000, 500))

# Convert annotation results to a data frame
annotated_dmr_df <- as.data.frame(dmr_annotation)

# Merge gene names into the annotation results
annotated_dmr_df <- merge(annotated_dmr_df, gene_info, by = "geneid", all.x = TRUE)

# Save annotated results
write.csv(annotated_dmr_df, file.path(output_dir, "output_eQTMoutput_annotations_CKD.csv"), row.names = FALSE)

# Print completion message
print("DMR annotation completed!")

```


