
---
title: "RNA-seq Processing Pipeline"
author: "Dr. Manasa Kalya"
date: "`r Sys.Date()`"
output: html_document
---

# RNA-seq Analysis Pipeline

This pipeline processes RNA-seq FASTQ files, performs quality control (QC), quantifies transcripts using Salmon, and generates gene-level count matrices.  
It combines shell-based and R-based steps into a unified workflow for reproducibility.

---

## 1. Sample Processing with Fastp and Salmon

This section assumes that `fastp`, `salmon`, and `samtools` are installed and accessible from the command line.

```{bash}
# Define paths (edit these to match your project layout)
INPUT_DIR="./data/fastq/"
OUTPUT_DIR="./output/salmon_results/"
SALMON_INDEX="./data/reference/salmon_index"

mkdir -p "$OUTPUT_DIR"

SAMPLES=("Sample1" "Sample2")  # Generalized sample names
LANES=("L001" "L002")

for SAMPLE in "${SAMPLES[@]}"; do
  SAMPLE_OUT="${OUTPUT_DIR}/${SAMPLE}"
  mkdir -p "$SAMPLE_OUT"

  for LANE in "${LANES[@]}"; do
    R1="${INPUT_DIR}/${SAMPLE}_${LANE}_R1.fastq.gz"
    R2="${INPUT_DIR}/${SAMPLE}_${LANE}_R2.fastq.gz"

    echo "Running Fastp for ${SAMPLE}_${LANE}..."
    fastp -i "$R1" -I "$R2" \
          -o "${SAMPLE_OUT}/${SAMPLE}_${LANE}_R1_trimmed.fastq.gz" \
          -O "${SAMPLE_OUT}/${SAMPLE}_${LANE}_R2_trimmed.fastq.gz" \
          --detect_adapter_for_pe -l 25

    echo "Running Salmon for ${SAMPLE}_${LANE}..."
    salmon quant -i "$SALMON_INDEX" -l A \
        -1 "${SAMPLE_OUT}/${SAMPLE}_${LANE}_R1_trimmed.fastq.gz" \
        -2 "${SAMPLE_OUT}/${SAMPLE}_${LANE}_R2_trimmed.fastq.gz" \
        --validateMappings --gcBias \
        -o "${SAMPLE_OUT}/${SAMPLE}_${LANE}_quant"
  done
done
```

---

## 2. Create Transcript-to-Gene Mapping

```{r tx2gene-mapping, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)

gtf_file <- "./data/reference/gencode.v42.annotation.gtf.gz"
gtf_data <- read_tsv(gtf_file, comment = "#", col_names = FALSE)

colnames(gtf_data) <- c("seqname", "source", "feature", "start", "end", "score",
                        "strand", "frame", "attribute")

transcripts_data <- gtf_data %>%
  filter(feature == "transcript") %>%
  mutate(
    gene_id = gsub('.*gene_id "([^"]+)".*', '\1', attribute),
    transcript_id = gsub('.*transcript_id "([^"]+)".*', '\1', attribute)
  ) %>%
  select(gene_id, transcript_id) %>%
  distinct() %>%
  na.omit()

write_tsv(transcripts_data, "./output/transcript_to_gene_mapping.tsv")
```

---

## 3. Summarize Transcript-Level Quantification to Gene-Level Counts

```{r summarize-counts, message=FALSE}
library(tximport)

samples <- c("Sample1", "Sample2")
tx2gene <- read_tsv("./output/transcript_to_gene_mapping.tsv")
tx2gene_df <- tx2gene[, c("transcript_id", "gene_id")]

quant_files <- file.path("./output/salmon_results", samples, paste0(samples, "_L001_quant", "/quant.sf"))

names(quant_files) <- samples

txi <- tximport(files = quant_files, type = "salmon", tx2gene = tx2gene_df, ignoreAfterBar = TRUE)
counts_matrix <- txi$counts

write.csv(counts_matrix, "./output/gene_counts_matrix.csv")
```

---

## 4. Merge Count Matrices and Generate Final Expression Table

```{r merge-matrices, message=FALSE}
library(readr)
library(dplyr)

df <- read.csv("./output/gene_counts_matrix.csv", row.names = 1)
df <- tibble::rownames_to_column(df, var = "gene_id")
write.csv(df, "./output/final_gene_expression_matrix.csv", row.names = FALSE)
```

---

# ✅ Done!

You now have a gene-level count matrix ready for downstream analysis like normalization, DESeq2, or visualization.
